---
title: "DATA 608 Assignment 1"
author: "Salma Elshahawy"
date: "August 28, 2020"
output:
  html_document: default
  html_notebook: default
---

**Principles of Data Visualization and Introduction to ggplot2**

I have provided you with data about the 5,000 fastest growing companies in the US, as compiled by Inc. magazine. lets read this in:

```{r}

inc <- read.csv("https://raw.githubusercontent.com/charleyferrari/CUNY_DATA_608/master/module1/Data/inc5000_data.csv", header= TRUE)

```

And lets preview this data:

```{r}

head(inc)

```
```{r}

summary(inc)

```

Think a bit on what these summaries mean. Use the space below to add some more relevant non-visual exploratory information you think helps you understand this data:

```{r}
library(dplyr)
# list types for each attribute
sapply(inc, class)
```

```{r}
# Insert your code here, create more chunks as necessary
sd(inc$Growth_Rate)
sd(inc$Revenue)
sd(inc$Employees, na.rm = TRUE)#A few companies have missing employee counts
```
```{r}
#We can also do IQR in case the data is skewed
IQR(inc$Growth_Rate)
IQR(inc$Revenue)
IQR(inc$Employees, na.rm = TRUE)
```

```{r}
#Revenue has quite a large range. I used a base 10 logrithm to compress that scale.
#I used mutate from tidyr to make these new calculations part of the dataframe for later use.
inc <- inc %>% mutate(log_Rev = log10(Revenue))
inc$log_Rev %>% summary()
```

```{r}
# I also like doing ratios, in this case normalizing businesses of different sizes to see how much revenue or growth is generated per employee. 
inc <- inc %>% 
  mutate(rev_per_empl = Revenue/Employees)
inc$rev_per_empl %>% 
  summary() 
inc <- inc %>% 
  mutate(grw_per_empl = Growth_Rate/Employees)
inc$grw_per_empl %>% 
  summary()
inc <- inc %>% 
  mutate(log_rev_per_grw = log10(Revenue/Growth_Rate))
inc$log_rev_per_grw %>% 
  summary()
```

```{r}
#  growth rate by industry
gr_industry<-inc %>% 
  select (Industry,Growth_Rate) %>% 
  group_by(Industry) %>% 
  summarise(avg_gr= mean(Growth_Rate)) %>% 
  arrange(desc(avg_gr))
head(gr_industry)
```

```{r}
#  growth rate by city

gr_city<-inc %>% 
  select (City,Growth_Rate) %>% 
  group_by(City) %>% 
  summarise(avg_gr= mean(Growth_Rate)) %>% 
  arrange(desc(avg_gr))
head(gr_city)
```


## Question 1

Create a graph that shows the distribution of companies in the dataset by State (ie how many are in each state). There are a lot of States, so consider which axis you should use. This visualization is ultimately going to be consumed on a 'portrait' oriented screen (ie taller than wide), which should further guide your layout choices.

```{r fig.height=10, fig.width=10}
library(ggplot2)
q1_data<-inc %>% 
  select (Name,State) %>% 
  group_by(State) %>% 
  dplyr::summarise(company_count = n_distinct(Name)) %>% 
  arrange(desc(company_count))

q1<-ggplot(q1_data, aes(x=reorder(State,company_count), y=company_count)) +
  geom_bar(stat="identity")+
  geom_col(aes(fill = company_count)) + 
  geom_point(size=0.5, colour = "red") +
  scale_fill_gradient2(low = "white", high = "red") + 
  theme_bw()+
  coord_flip() + 
  theme(text = element_text(size = 9, color = "black")) +
  ggtitle ("Number Of Fastest Growing Companies By State") + ylab("Number of Companies") +
  theme(axis.title.y=element_blank()) + 
  theme(legend.position="none")

q1
```


## Quesiton 2

Lets dig in on the state with the 3rd most companies in the data set. Imagine you work for the state and are interested in how many people are employed by companies in different industries. Create a plot that shows the average and/or median employment by industry for companies in this state (only use cases with full data, use R's `complete.cases()` function.) In addition to this, your graph should show how variable the ranges are, and you should deal with outliers.

```{r}

# Answer Question 2 here
ny_data <- inc  %>% 
  filter(State == 'NY', complete.cases(.)) %>% 
  arrange(Industry) %>% select(Industry, Employees) 
ny_data <- ny_data %>% group_by(Industry) %>% 
  filter(!(abs(Employees - median(Employees)) > 1.5*IQR(Employees)))# Using 1.5xIQR as the outlier limit
ny_data
#The 1.5xIQR rule reduced the number of negative error bars better than the 2xstd dev rule.
ind_means <- ny_data %>% 
  group_by(Industry) %>% 
  summarise(mean_emp = mean(Employees), emp_sd = sd(Employees))
ind_means$emp_sd[is.na(ind_means$emp_sd)] <- 0
ind_means
ggplot(ind_means, aes(x=reorder(Industry, mean_emp),y=mean_emp)) +
  geom_bar(stat='identity', color = 'black', fill='lightgray') +
  geom_errorbar(aes(ymin = mean_emp - emp_sd, ymax = mean_emp + emp_sd), width=0.2) +
  theme(legend.position="none") +
  ylab('Mean Employees')+ xlab('Industry')+ 
  coord_flip() +
  theme_classic()

```


## Question 3

Now imagine you work for an investor and want to see which industries generate the most revenue per employee. Create a chart that makes this information clear. Once again, the distribution per industry should be shown.

```{r}

# Answer Question 3 here
q3_data<-inc %>% 
  select (Revenue, Industry, Employees) %>% 
  group_by(Industry) %>%
  summarise(total_revenue = sum(Revenue), total_employee = sum(Employees)) %>%
  mutate(revenue_employee = total_revenue / total_employee/1000) %>% 
  arrange (revenue_employee)

q3_data <- q3_data[complete.cases(q3_data$Industry), ]
q3_data <- q3_data[complete.cases(q3_data$total_employee), ]

q3<-ggplot(q3_data, aes(x=reorder(Industry, revenue_employee), y=revenue_employee)) +
  geom_bar(stat="identity")+
  theme_bw()+
  geom_col(aes(fill = revenue_employee)) + 
  geom_point(size=0.5, colour = "red") +
  scale_fill_gradient2(low = "white", high = "red") + 
  coord_flip() + 
  ggtitle ("Revenue Generated Per Employee By Industry") + ylab("Revenue Per Employee, in thousands") + 
  theme(legend.position="none") +
  theme(axis.title.y=element_blank())+
  theme(text = element_text(size = 8, color = "black"))

q3
```


